### 2017-11-09 papers/cvpr2018/supplemental/skeletonization.tex
Add supplemental material template.

\section{Skeletonization}

more details regarding the skeletonization approach, which I believe was extra fast?

\subsection{Parameters}

\subsection{Performance Tests}