### 2017-11-10 papers/cvpr2018/paper.tex
added some figures

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{siunitx}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{0446} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Graph-based Neuron Agglomeration using 3D Skeletons}

%Robust Agglomeration using 3D Skeletonization of Labeled for Connectomics}

\author{Brian Matejek\\
Harvard University\\
Cambridge, MA 02138, USA\\
{\tt\small bmatejek@seas.harvard.edu}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
%\and
%Second Author\\
%Institution2\\
%First line of institution2 address\\
%{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}

\begin{figure}[t]\footnotesize
	\vspace{-2.04in}
	\newcommand{\FigureWidth}{0.23\textwidth}
	\newcommand{\FigureShift}{\hspace{-0.42cm}}
	\newcommand{\figureShift}{\hspace{-0.3cm}}
	\begin{minipage}{\textwidth}
		\begin{center}
			\begin{tabular}{cccc}
				\FigureShift \includegraphics[width=\FigureWidth]{figures/schema/teaser-image.png} \includegraphics[width=\FigureWidth]{figures/schema/teaser-segmentation.png} &
				\FigureShift \includegraphics[width=\FigureWidth]{figures/schema/pre-multicut.png} &
				\FigureShift \includegraphics[width=\FigureWidth]{figures/schema/multicut-graph.png}&
				\FigureShift \includegraphics[width=\FigureWidth]{figures/schema/post-multicut.png}\\
				\figureShift (a) Input & \FigureShift (b)  Our results & \FigureShift(c) Dark channel of (a) & \FigureShift (d) Dark channel of (b)\\
			\end{tabular}
		\end{center}
		\vspace{-3mm}
		\caption{Deblurring result on a challenging low-light image. The blur process makes the dark channel of the blurred image less sparse~(c). Enforcing sparsity on the dark channel of the recovered image favors clean images over blurred ones.}
		\label{fig:teaser}
		%
	\end{minipage}
\end{figure}

%%%%%%%%% ABSTRACT
\begin{abstract}
Advancements in electron microscopy image acquisition have created massive connectomics datasets which are petabytes in size and make manual segmentation not feasible. 
Proposed methods for automatic segmentations generate super-pixel approximations of neural membranes followed by agglomeration strategies to create full neuron reconstructions. 
However, pixel-based segmentation results widely neglect global geometric properties.
We generate skeletons of membrane labels to approximate neural pathways in 3D. This allows us to efficiently apply graph-based optimization strategies and to train automatic classifiers for shape description against manually labeled ground truth. Our classifiers detect feasible and impossible neural pathways by looking at geometric properties alone, and are able to improve the segmentation labels accordingly. We demonstrate the performance of our classifiers on multiple real-world connectomics datasets with average variation of information improvement of $X\times$. We discuss our findings and provide insights to learning shape features for neuron agglomeration.
\end{abstract}

%%%%%%%%% BODY TEXT
\input{0_introduction}

\input{1_related_work}

\input{2_method}

\input{3_evaluation}

\input{4_results}

\input{5_conclusions}

{\small
\bibliographystyle{ieee}
\bibliography{paper}}

\end{document}