### 2017-11-13 papers/cvpr2018/0_introduction.tex
Changed sections 1 and 2 and some other errors such as duplicate bibtex.

\section{Introduction}

The field of connectomics is concerned with reconstructing the wiring diagram of the brain at nanometer resolutions. 
Recent advancements in image acquisition using multi-beam serial-section electron microscopy (sSEM) has allowed researchers to produce terabytes of electron microscopy (EM) image data every hour~\cite{hildebrand2017whole}.
Neuroscientists believe that reconstructing an entire mammalian brain at fine resolution will enable new insights into the workings of the brain~\cite{kasthuri2015saturated,haehn2017scalable}. 
These observations will allow for new advancements in neuromedicine and artificial intelligence.

Segmenting the image stack into a label volume is a substantial component of this reconstruction process.
In a label volume, every voxel receives a segment label corresponding to a unique neuron. 
Voxels with the same label belong to the same neuron.
It is not feasible for domain experts to manually segment the vast amount of 3-D image data to model an entire brain~\cite{haehn2014design}.
Here we introduce a scalable, top-down segmentation algorithm that also leverages local information.
Partitioning globally allows us to apply domain-specific constraints to the segmentation task.

\begin{figure*}[t!]
	\centering
	\includegraphics[width=0.22\linewidth]{./figures/pipeline/image.png}
	\hspace{0.025\linewidth}
	\includegraphics[width=0.22\linewidth]{./figures/pipeline/affinities.png}
	\hspace{0.025\linewidth}
	\includegraphics[width=0.22\linewidth]{./figures/pipeline/watershed.png}
	\hspace{0.025\linewidth}
	\includegraphics[width=0.22\linewidth]{./figures/pipeline/neuroproof.png}
	\caption{An overview of an existing pipeline in segmentation for connectomics. From left to right: the original EM image data, the output of a CNN predicting voxel affinities, clustering of the affinities using a watershed algorithm, an agglomeration of the supervoxels into larger segments.}
	\label{fig:pipeline}
\end{figure*}

A significant amount of research focuses on automatic reconstruction of the neurons in EM images~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image}. %TODO DOUBLE CHECK REFERENCES
All of these algorithms extract neuronal processes through the 3-D volumes using the raw image data as input.
Oftentimes, convolutional neural networks (CNNs) predict membrane probabilities or affinities between voxels and apply simple thresholds to agglomerate the voxels into clusters~\cite{lee2015recursive,ronneberger2015u}.
These \textit{pixel-based} algorithms produce excellent results but currently fall short of complete reconstructions with error rates of approximately $X\%$. %TODO ADD ERROR RATES

Researchers address the failures of these pixel-based algorithms by training random-forest classifiers to agglomerate an oversegmentation of voxels~\cite{nunez2014graph,10.1371/journal.pone.0125825}.
These classifiers take the output of the pixel-based algorithms as input and generate high-level statistics such as affinity distributions between regions. 
Presently, these methods use hand-designed features despite the evidence that machine-learned features perform better~\cite{bogovic2013learned}. 
These \textit{region-based} algorithms outperform the pixel-based algorithms but do not provide the accuracy needed for large scale reconstructions of the brain.

We present a \textit{segment-based} strategy that builds on the outputs of region-based methods (Figure \ref{fig:teaser}).
For training a classifier based on a CNN, we only take as input the segmentation from a region-based agglomeration algorithm. 
This renders our classifier independent of image resolution and characteristics such as anisotropy and for instance, allows the application to isotropic data without retraining.
This is a huge plus considering the sheer amount of expert-hours needed to generate ground truth for these massive datasets.
We extract a graph representation from the input segmentation.
By simplifying the 3-D segments by their skeletons, we can quickly generate the graph nodes and edges. 
We train a 3-D CNN to populate the graph edge weights with bottom-up local information. 
We enforce domain-specific global constraints, such as the underlying biology, using a graph-based partitioning heuristic.
In this we can leverage both local and global information to produce a more accurate segmentation.

Our approach creates a level of abstraction over existing region-based agglomeration strategies. 
We take the output of these methods as input and extract a graph structure from that.
This allows us to enforce global restrictions on our output that closely follow the underlying biology of these connectomics datasets such as limiting the allowed angles of neuron branching. 
In addition we use local information such as shape priors to weight the edges in our graph.
This dual approach of assessing local decisions in a global framework produces improvements over the existing region-based methods.