### 2017-11-13 papers/cvpr2018/0_introduction.tex
added HP comments

% !TEX root =  paper.tex
\section{Introduction}

The field of connectomics is concerned with reconstructing the wiring diagram of the brain at nanometer resolutions to enable new insights into the workings of the brain~\cite{kasthuri2015saturated,haehn2017scalable}. Recent advancements in image acquisition using multi-beam serial-section electron microscopy (sSEM) have allowed researchers to produce terabytes of image data every hour~\cite{hildebrand2017whole}. It is not feasible for domain experts to manually reconstruct thousands or millions of neurons from this vast amount of image data~\cite{haehn2014design}. State-of-the-art automatic reconstruction approaches use pixel-based segmentation with convolutional neural networks (CNNs) followed by agglomeration strategies~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,lee2015recursive,ronneberger2015u}. %TODO DOUBLE CHECK REFERENCES
These \textit{bottom-up pixel-based} algorithms produce excellent results but still fall short of acceptable error rates for large volumes.

\begin{figure*}[t!]
	\centering
	(a) \includegraphics[width=0.18\linewidth]{./figures/pipeline/image.png}
	\hspace{0.025\linewidth}
	(b) \includegraphics[width=0.18\linewidth]{./figures/pipeline/affinities.png}
	\hspace{0.025\linewidth}
	(c) \includegraphics[width=0.18\linewidth]{./figures/pipeline/watershed.png}
	\hspace{0.025\linewidth}
	(d) \includegraphics[width=0.18\linewidth]{./figures/pipeline/neuroproof.png}
	\caption{An overview of a pixel-based connectomics segmentation pipeline. (a) Original EM image data (b) Output of a CNN predicting voxel affinities (c) Clustering of the affinities using a watershed algorithm (d) Agglomeration of the super-voxels into larger segments.}
	\label{fig:pipeline}
\end{figure*}

%Researchers address the failures of these pixel-based algorithms by training random-forest classifiers to agglomerate an oversegmentation of voxels~\cite{nunez2014graph,10.1371/journal.pone.0125825}.
%These classifiers take the output of the pixel-based algorithms as input and generate high-level statistics such as affinity distributions between regions.
%Presently, these methods use hand-designed features despite the evidence that machine-learned features perform better~\cite{bogovic2013learned}.
%These \textit{region-based} algorithms outperform the pixel-based algorithms but do not provide the accuracy needed for large scale reconstructions of the brain.

%Here we introduce a scalable, top-down segmentation algorithm that also leverages local information.
%Partitioning globally allows us to apply domain-specific constraints to the segmentation task.
%We extract a graph representation from the input segmentation.
%By simplifying the 3-D segments by their skeletons, we can quickly generate the graph nodes and edges.

We present a \textit{top-down graph-based} method that builds on the outputs of bottom-up pixel-based segmentation approaches (Fig.~\ref{fig:teaser}a). We first extract 3D skeletons from the input segmentation (Fig.~\ref{fig:teaser}b) and generate a simplified 3D graph (Fig.~\ref{fig:teaser}c). We train a 3D CNN classifier on the agglomerated regions in the segmentation data to detect errors. During test time, we run the classifier to populate the graph edge weights with error probabilities. We then use a graph optimization algorithm to partition the graph into the final reconstruction by enforcing domain-specific global constraints from the underlying biology (Fig.~\ref{fig:teaser}d).

Our approach operates at a level of abstraction above existing pixel-based methods. This allows us to leverage both local and global information to produce more accurate reconstructions. Because it uses agglomerated regions our classifier is independent of image resolution and acquisition parameters, enabling its application to isotropic and anisotropic image data without retraining. Using the 3D graph of the segmentation allows us to enforce global constraints on the reconstruction that closely follow the underlying biology, such as limiting the angles of neuron branchings to the biologically plausible range. We can also use local information such as shape priors to weight the edges in our graph. This dual approach of assessing local decisions in a global context yields accuracy improvements over the existing reconstruction methods.

This work makes the following following contributions: (1) a novel top-down graph-based method for neural reconstruction of connectomics data; (2) a region-based CNN classifier to detect errors under global constraints; (3) an empirical evaluation of our method on several connectomics datasets; (4) our method yields improved performance over the state-of-the-art on X out of Y datasets, and competitive results on the remaining data. On average, we improve the Variation of Information (VI) metric across all datasets by X percent without drastically increasing the running time.