### 2017-11-15 papers/cvpr2018/5_conclusions.tex

% !TEX root =  paper.tex
\section{Conclusions}

We present a novel method for reconstructing neuronal processes in connectomics datasets. We extend existing pixel-based reconstruction strategies using a morphological 3D skeleton graph and show significant accuracy improvement on two datasets from two different species. Our approach enables enforcing domain-constraints on the resulting segmentation. We believe this approach extends beyond connectomics to other domains with different constraints.
\hp{add more future work}