### 2017-11-15 papers/cvpr2018/5_conclusions.tex

% !TEX root =  paper.tex
\section{Conclusions}

We present a novel method for reconstructing neuronal processes in connectomics datasets. 
We extend existing region-based reconstruction strategies using a morphological 3D skeleton graph and show significant accuracy improvement on two datasets from two different species. 
Our approach enables enforcing domain-constraints on the resulting segmentation. 
We believe this approach extends beyond connectomics to other domains with different constraints.

There is significant room for additional research related to this topic.
We can still extract a lot more information from the skeletons of each segment.
Currently these skeletons identify potential merge locations.
However, these representations can be used to identify improperly merged segments.
Additionally, we can further improve our results by tweaking the partitioning cost functions to more closely match the underlying biology.
One such extension might take the outputs of a synaptic detector to enforce the constraint that a given segment only has post- or pre-synaptic connections.