### 2017-11-22 papers/cvpr2018/supplemental/skeletonization.tex
updated supplemental

\section{Skeletonization}

more details regarding the skeletonization approach, which I believe was extra fast?

\subsection{Node Pruning}

\subsection{Edge Pruning}

\subsection{CNN Parameters}

We considered the following sets of parameters for the neural network and ran a brute-force search over the parameters.

\begin{table}
	\footnotesize
	\begin{tabular}{c c} \hline
		\textbf{Parameter} & \textbf{Considered Options} \\ \hline
		Optimizer & \textbf{SGD with Nesterov}, Adam \\
		Loss Function & Binary Cross Entropy, \textbf{Mean Squared Error} \\
		Output Size & 3, 4, 5, \textbf{6}, 7 \\
		Batch Normalization & On, \textbf{Off} \\ \hline
	\end{tabular}
\end{table}

We also consider networks with four layers of double convolutions and increased filter sizes $(16\times32\times64)$ but both these configurations have two many parameters to learn with such small datasets.

\subsection{Performance Tests}

We use a XX GHz Processor with a PASCAL TitanX 1080 GPU with 64GB of RAM for all of our performance tests. 
Inference for the CNN takes an average of $6.2$ seconds per $1000$ examples on our network of size