### 2017-11-22 papers/cvpr2018/supplemental/performance.tex

\section{Running Times}

\subsection{System}

All performance experiments ran on an Intel i7-6800K CPU 3.4 0GHz with a PASCAL Titan X GPU. 
All code is written in Python and will be freely available upon the paper decision.
We use the Keras deep learning library for our neural networks with Theano backend with cuDNN 7 acceleration for cuda 8.0.

\subsection{Performance}

The network trained for $34$ epochs with each epoch taking approximately $1200$ seconds to complete.
The following tables show the total running time for each dataset for graph creation, CNN inference, and multicut partitioning. 
In all instances the time for these three steps was less than 100 seconds.
Currently extracting the skeletons with the TEASER algorithm is the bottleneck to the process.
The skeleton for each segment takes around $20-30$ seconds to generate. 
However, the skeleton process is parallelizable and each segment can be processed independently.
Faster skeleton extraction is an important area for future research.

\begin{table}[t]
	\centering
	\small
	\begin{tabular}{c c c } \hline
		& Kasthuri Vol. 1 & Kasthuri Vol. 2 \\ \hline
		Graph Generation & 34.76s & 35.90s \\
		Inference & 38.47s & 37.13s \\
		Multicut & 22.94s & 22.90s \\ \hline
		Total Time & 96.17s & 95.93s \\ \hline
	\end{tabular}
	\begin{tabular}{c c c } \hline
		& FlyEM Vol. 1 & FlyEM Vol. 2 \\ \hline
		Graph Generation & 46.52s & 48.97s \\
		Inference & 12.55s & 10.91s \\
		Multicut & 30.30s & 30.14s \\ \hline
		Total Time & 89.37s & 90.02s \\ \hline
	\end{tabular}
\end{table}