### 2018-03-09 papers/eccv2018/supplemental/classifier.tex

\section{Merge Classification}

\subsection{Architecture}

\subsection{Training Augmentation}

\subsection{Testing Augmentation}