### 2018-03-12 papers/eccv2018/supplemental/skeletonization.tex

\section{Skeletonization}

\subsection{Volume Downsampling}

\subsection{Node Generation}

\subsection{Edge Generation}