### 2018-04-27 papers/quals2018/paper.tex
Changed format and removed old text

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{subfig}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{siunitx}
\usepackage{float}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{0446} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
%

\cvprfinalcopy

%% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

	%%%%%%%%% TITLE
	\title{Segmentation of Electron Microscopy Images for Connectomics}

	\author{Brian Matejek\\
		Advisor: Hanspeter Pfister \\
	Harvard University\\
	{\tt\small bmatejek@seas.harvard.edu}
	%% For a paper whose authors are all at the same institution,
	%% omit the following lines up until the closing ``}''.
	%% Additional authors and addresses can be added with ``\and'',
	%% just like the second author.
	%% To save space, use either the email address or home page, not both
	%%\and
	%%Second Author\\
	%%Institution2\\
	%%First line of institution2 address\\
	%%{\tt\small secondauthor@i2.org}
	}

	\maketitle
	%\thispagestyle{empty}
	%

	\input{0_introduction}

	\input{1_related_work}

	\input{2_method}

	\input{3_results}

	\input{4_proposed}

	\input{5_conclusions}

	{\small
		\bibliographystyle{ieee}
		\bibliography{paper}}

\end{document}