### 2018-04-27 papers/quals2018/0_introduction.tex
updated proposal

\section{Introduction}

The field of connectomics is concerned with the reconstruction of the wiring diagram of the mammalian brain.
In order to achieve this ambitious goal, we need images at nanometer resolution to see the structure of the neurons and the synaptic connections between them.
Recent advances in electron microscopy acquisition techniques enable a throughput of a terabyte of image data every hour~\cite{richard2016imaging}.
Manual reconstruction of this image data is simply infeasible and requires automatic segmentation methods.
These automatic methods label every voxel of the image with a 64-bit integer value where two voxels have the same label only if they belong to the same neuron.
As the field progresses, our goal is to reconstruct a cubic millimeter of brain which will result in over 18 petabytes of segmentation data alone~\cite{suissa2016automatic}.

Three major challenges arise as the image datasets scale to a petabyte and more. 
First, the segmentation methods must be accurate enough to extract the proper connections between neurons.
In particular, the dendritic spines \TODO{check,CITE} which contain the postsynaptic density can be only \TODO{XX} nanometers thick.
Second, automatic segmentation must be fast -- ideally matching the throughput of the electron microscope~\cite{haehn2017scalable}.
The current state-of-the-art automatic reconstruction pipelines are either too slow for such large datasets or too inaccurate at scale.
Third, the label volumes require $8\times$ the number of bytes as the already massive image volumes.
Storing such data is expensive and transferring the data is slow. 
To cut costs and delays, we need compression methods to reduce data sizes.