### 2018-05-24 papers/eccv2018/rebuttal/paper.tex
skeleton: allow locations close to border

% last updated in April 2002 by Antje Endemann
% Based on CVPR 07 and LNCS, with modifications by DAF, AZ and elle, 2008 and AA, 2010, and CC, 2011; TT, 2014; AAS, 2016

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
%\usepackage{ruler}
\usepackage{color}
\usepackage{subfig}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{siunitx}
\usepackage{float}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}

\begin{document}
% \renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
% \renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}}
% \linenumbers
\pagestyle{headings}
\mainmatter

%\titlerunning{Biologically-Constrained Region Merging for Connectome Reconstruction}
%\authorrunning{B. Matejek et al.}
\title{Biologically-Constrained Region Merging for Connectome Reconstruction}

%\author{Brian Matejek, Daniel Haehn, Donglai Wei, Toufiq Parag, Hanspeter Pfister}
%\institute{Harvard University, Cambridge, MA 02138, USA\\
%	\email{bmatejek,haehn,donglai,paragt,pfister@seas.harvard.edu}}

\newcommand\TODO[1]{\textcolor{red}{#1}}

\maketitle

Hello World

\bibliographystyle{splncs}
\bibliography{egbib}

\end{document}