### 2018-08-13 papers/cvpr2019/paper.tex
papers: added framework for CVPR 2019 submission

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{subfig}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{siunitx}
\usepackage{float}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{0446} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
%
\newcommand{\FIX}[1]{\textcolor{red}{#1}}
\newcommand{\hp}[1]{\textcolor{red}{HP: #1}}
\newcommand{\donglai}[1]{\textcolor{blue}{[Donglai: #1]}}

%% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

	%%%%%%%%% TITLE
	\title{Biologically-Constrained Region Merging for Connectome Reconstruction}

	%\author{Brian Matejek\\
	%Harvard University\\
	%Cambridge, MA 02138, USA\\
	%{\tt\small bmatejek@seas.harvard.edu}
	%% For a paper whose authors are all at the same institution,
	%% omit the following lines up until the closing ``}''.
	%% Additional authors and addresses can be added with ``\and'',
	%% just like the second author.
	%% To save space, use either the email address or home page, not both
	%%\and
	%%Second Author\\
	%%Institution2\\
	%%First line of institution2 address\\
	%%{\tt\small secondauthor@i2.org}
	%}

	\maketitle
	%\thispagestyle{empty}
	%

	%%%%%%%%% ABSTRACT
	\begin{abstract}
	Current state-of-the-art connectome reconstruction pipelines have two major steps: initial pixel-based segmentation with affinity prediction and watershed transform, and refined segmentation by merging over-segmented regions.
	For the region merging, most methods neglect geometric and topological constraints that are inherent in the graph structure of neural wiring diagrams.
	In this work, we use a biologically-constrained graph-based method to better reconstruct neural pathways.
	Given the initial segmentation, we first create a graph with nodes corresponding to segment labels and edges identifying potential locations of split errors based on the skeleton structure of the segmentation.
	Next, we train a convolutional neural network to capture the geometric properties of segment connections and produce edge weights as the probability of merging two nodes.
	Finally, we formulate the region merging as a lifted multicut partitioning problem with topological constraints to avoid cycles.
	We demonstrate the performance of our approach on multiple real-world connectomics datasets with an average variation of information improvement of 8.2\%.
	\end{abstract}
	%%%%%%%%	% BODY TEXT

	\input{0_introduction}

	\input{1_related_work}

	\input{2_method}

	\input{3_evaluation}

	\input{4_results}

	\input{5_conclusions}

	{\small
		\bibliographystyle{ieee}
		\bibliography{paper}}

\end{document}