### 2018-08-13 papers/cvpr2019/supplemental/skeleton-generation.tex

% !TEX root =  supplemental.tex
\section{Skeletonization}

There are three steps in our skeleton generation pipeline: downsampling, skeletonization, and upsampling. 
Here we explore the advantages and disadvantages of each step both qualitatively by visualing inspecting results and quantitatively by publishing a novel ``Skeleton Benchmark''. 
Lastly, we compare three widely used skeletonization algorithms against our benchmark to determine the relative efficiencies and accuracies on segmented neuronal data.

\subsection{Skeleton Pipeline}

First, we downsample the original image to a resolution of $t_{res}\si{\nm^3}$ where $t_{res}$ is a tunable parameter. 
We experiment with values of $t_{res}$ ranging from $30$ to $200$ and display the results in Sec.~\ref{skeleton-results}.
In general, topologically preserving downsampling is impossible~\cite{kraus2001topology} as holes can collapse when the width of a gap is smaller than the downsampling ratio.
This is particularly common on dendrites where multiple spines protrude closely.

There are two major benefits to downsampling the data before running the skeletonization algorithm.
First, the overall skeletonization running time greatly decreases as a function of downsample resolution.
Two of our baseline skeleton algorithms have running times linear in the number of voxels in the input segmentation.
The third uses Dijkstra's algorithm to find globally distance voxels to form branches of the skeleton ``tree''.
Therefore, the computational complexity is $\mathcal{O}(n \log{n})$ since the number of edges in a structured voxel grid is $\mathcal{O}(n)$.
Sec.~\ref{skeleton-results} further explores the running time of the skeleton algorithms as a function of downsample rate.
Second, downsampling reduces the number of spurious endpoints.
Our input segmentations often have noisy boundaries with small spikes protruding from the surface that do not represent larger dendritic spines.
Downsampling the input segmentations has a smoothing effect that eliminates these small bumps while

The second step is to run the actual skeletonization algorithm

The last step of the algorithm ``upsamples'' the skeleton to the resolution of the original image.
If we downsample in $x$, $y$, and $z$ by $down_x$, $down_y$, and $down_z$ respectively, we cannot merely take a skeleton joint at $(u, v, w)$ and identify $(down_x * u, down_y * v, down_z * w)$ as the upsampled joint since this point may not fall within the original label boundary.
It is during this process that we also identify which skeleton joints are also endpoints.

\subsection{Skeleton Benchmark}

We created a skeleton benchmark to evaluate three different skeletonization approaches with varying downsampling parameters.

\subsection{Skeleton Results}
\label{skeleton-results}

\paragraph{Endpoint Accuracy.}

\paragraph{Running Times.}